---
title: "ETL Transform Census Data"
author: "Steven Macapagal"
date: "2023-09-07"
output: html_document
---

## Setup

- Last modified: 2025-05-05
- Last deployed (development): 2023-11-11
- Last deployed (production): 
- Inputs: aline.orr/acs_data
- Outputs: sf_dot_density_mvp

```{r setup_doc}
knitr::opts_chunk$set(echo = TRUE)

# allows pins to work from RSConnect 
httr::set_config(httr::config(ssl_verifypeer = FALSE, ssl_verifyhost = FALSE))
```

### Load libraries

```{r setup_libraries}
library(tidyverse)
library(terra)       # engine used to create dot densities
library(sf)
library(sfarrow)
library(qs)
library(tidycensus)  # as_dot_density() wrapper
library(pins)
```

### Read pins

```{r read_pins}
board <- board_connect(auth = "envvar") # connect to Posit Connect board
sf_acs <- pin_read(board, "aline.orr/acs_data") # read this pin
```

```{r}
# working_dir <- here::here("app")
sf_acs <- st_read_feather(here::here("data/sf_acs_all_2023.feather"))
```


```{r}
df_counties_tx <- tribble(
  ~ state, ~ region, ~ county,
  
  "TX", "Austin", "Travis County",
  "TX", "Austin", "Hays County",
  # "TX", "Austin", "Bastrop County",
  "TX", "Austin", "Williamson County",
  # "TX", "Austin", "Caldwell County",
  
  "TX", "Coastal Bend", "Nueces County",

  "TX", "El Paso", "El Paso County",
  
  "TX", "Greater Houston Area", "Harris County",
  # "TX", "Greater Houston Area", "Montgomery County",

  "TX", "Permian Basin", "Midland County",
  "TX", "Permian Basin", "Ector County",
  
  "TX", "Rio Grande Valley", "Cameron County",
  "TX", "Rio Grande Valley", "Hidalgo County",
  "TX", "Rio Grande Valley", "Starr County",
  
  "TX", "San Antonio", "Bexar County",
  # "TX", "San Antonio", "Comal County",
  
  "TX", "Tarrant County", "Tarrant County",
  # "TX", "Tarrant County", "Parker County",
  # "TX", "Tarrant County", "Burleson County",
  "TX", "Tarrant County", "Dallas County"
)

df_counties_ips <- tribble(
  ~ state, ~ region, ~ county,
  
  "OH", "Cincinnati", "Hamilton County",
  
  "FL", "Jacksonville", "Duval County",
  "FL", "Tampa", "Hillsborough County",
  "FL", "Tampa", "Polk County",
  "FL", "Orlando", "Orange County",
  
  "LA", "Southern Louisiana", "East Baton Rouge Parish",
  "LA", "Southern Louisiana", "Orleans Parish"
)

df_counties_all <- bind_rows(
  df_counties_tx,
  df_counties_ips
)
```

## Transform data

```{r dot_density_children_poverty}
# children in poverty ----
# time: 1452.64 sec

# sf_children_in_poverty <- data.frame(
#   table_name = character(),
#   table_short_name = character(),
#   table_number = character(),
#   state = character(),
#   region = character(),
#   county = character(),
#   geography = character(),
#   GEOID = character(),
#   variable = character(),
#   estimate = double(),
#   moe = double(),
#   geometry = list()
# )

sf_children_in_poverty <- sf_acs_all %>% # filter(county %in% df_counties_tx$county) %>%
  
  # get appropriate table rows
  filter(table_short_name == "Children in poverty") %>%
  filter(!st_is_empty(.)) %>% # filter out NA geometry to prep for dot density
  
  # convert to dot density
  tidycensus::as_dot_density(
    value = "estimate",
    values_per_dot = 20
  ) # convert MULTIPOLYGON to 20:1 POINT geometry


# st_write_feather(sf_children_in_poverty, glue::glue("{working_dir}/sf_children_poverty_dd_2023.feather"))
```

```{r}
# children by age group, block group ----
# time: 10300.86 sec

sf_children_by_age_group_block_group_tx <- sf_acs %>% filter(county %in% df_counties_tx$county) %>%
  
  # get appropriate table rows
  filter(
    table_short_name == "Children by age group and sex",
    geography == "block group"
  ) %>% # pull specific ACS table
  filter(!st_is_empty(.)) %>% # filter out NA geometry to prep for dot density
  
  # consolidate variable groups
  mutate(
    variable = case_when(
      str_detect(variable, "Under 5 years") ~ "Under 5 years",
      str_detect(variable, "5 to 9 years") ~ "5 to 9 years",
      str_detect(variable, "10 to 14 years") ~ "10 to 14 years",
      str_detect(variable, "15 to 17 years") ~ "15 to 17 years",
      TRUE ~ variable
    )
  ) %>%
  group_by(table_name, table_short_name, table_number, state, county, geography, GEOID, variable) %>%
  summarize(
    estimate = sum(estimate),
    moe = moe_sum(moe, estimate)
  ) %>%
  ungroup() %>%
  
  # convert to dot density
  tidycensus::as_dot_density(
    value = "estimate",
    values_per_dot = 20
  ) # convert MULTIPOLYGON to 20:1 POINT geometry

sf_children_by_age_group_block_group <- bind_rows(
  sf_children_by_age_group_block_group_tx,
  sf_children_by_age_group_block_group_ips
)
# st_write_feather(sf_children_by_age_group_block_group, glue::glue("{working_dir}/sf_children_by_age_group_block_group_dd_2023.feather"))
```


```{r}
# children by age group, tract ----

sf_children_by_age_group_tract <- sf_acs %>%
  
  # get appropriate table rows
  filter(
    table_short_name == "Children by age group",
    geography == "tract"
  ) %>% # pull specific ACS table
  filter(!st_is_empty(.)) %>% # filter out NA geometry to prep for dot density
  
  # convert to dot density
  tidycensus::as_dot_density(
    value = "estimate",
    values_per_dot = 20
  ) # convert MULTIPOLYGON to 20:1 POINT geometry

# st_write_feather(sf_children_by_age_group_tract, glue::glue("{working_dir}/sf_children_by_age_group_tract_dd_2023.feather"))
```


```{r}
# household income ----
# time: 8074.69 sec

sf_household_income_houston <- sf_acs_houston %>%
  
  # get appropriate table rows
  filter(table_short_name == "Household income") %>% # pull specific ACS table
  filter(
    variable != "Total",
    !st_is_empty(.)
  ) %>% # filter out NA geometry to prep for dot density
  
  # consolidate variable groups
  mutate(
    variable = case_when(
      variable %in% c("Less than $10,000", "$10,000 to $14,999", "$15,000 to $19,999") ~ "Less than $20,000",
      variable %in% c("$20,000 to $24,999", "$25,000 to $29,999") ~ "$20,000 to $29,999",
      variable %in% c("$30,000 to $34,999", "$35,000 to $39,999") ~ "$30,000 to $39,999",
      variable %in% c("$40,000 to $44,999", "$45,000 to $49,999") ~ "$40,000 to $49,999",
      variable %in% c("$100,000 to $124,999", "$125,000 to $149,999", "$150,000 to $199,999", "$200,000 or more") ~ "$100,000 or more",
      TRUE ~ variable
    )
  ) %>%
  group_by(table_name, table_short_name, table_number, state, county, geography, GEOID, variable) %>%
  summarize(
    estimate = sum(estimate),
    moe = moe_sum(moe, estimate)
  ) %>%
  ungroup() %>%
  
  # convert to dot density
  tidycensus::as_dot_density(
    value = "estimate",
    values_per_dot = 20
  ) # convert MULTIPOLYGON to 20:1 POINT geometry

sf_household_income_tarrant <- sf_acs_tarrant %>%
  
  # get appropriate table rows
  filter(table_short_name == "Household income") %>% # pull specific ACS table
  filter(
    variable != "Total",
    !st_is_empty(.)
  ) %>% # filter out NA geometry to prep for dot density
  
  # consolidate variable groups
  mutate(
    variable = case_when(
      variable %in% c("Less than $10,000", "$10,000 to $14,999", "$15,000 to $19,999") ~ "Less than $20,000",
      variable %in% c("$20,000 to $24,999", "$25,000 to $29,999") ~ "$20,000 to $29,999",
      variable %in% c("$30,000 to $34,999", "$35,000 to $39,999") ~ "$30,000 to $39,999",
      variable %in% c("$40,000 to $44,999", "$45,000 to $49,999") ~ "$40,000 to $49,999",
      variable %in% c("$100,000 to $124,999", "$125,000 to $149,999", "$150,000 to $199,999", "$200,000 or more") ~ "$100,000 or more",
      TRUE ~ variable
    )
  ) %>%
  group_by(table_name, table_short_name, table_number, state, county, geography, GEOID, variable) %>%
  summarize(
    estimate = sum(estimate),
    moe = moe_sum(moe, estimate)
  ) %>%
  ungroup() %>%
  
  # convert to dot density
  tidycensus::as_dot_density(
    value = "estimate",
    values_per_dot = 20
  ) # convert MULTIPOLYGON to 20:1 POINT geometry

sf_household_income_ctx <- sf_acs_ctx %>%
  
  # get appropriate table rows
  filter(table_short_name == "Household income") %>% # pull specific ACS table
  filter(
    variable != "Total",
    !st_is_empty(.)
  ) %>% # filter out NA geometry to prep for dot density
  
  # consolidate variable groups
  mutate(
    variable = case_when(
      variable %in% c("Less than $10,000", "$10,000 to $14,999", "$15,000 to $19,999") ~ "Less than $20,000",
      variable %in% c("$20,000 to $24,999", "$25,000 to $29,999") ~ "$20,000 to $29,999",
      variable %in% c("$30,000 to $34,999", "$35,000 to $39,999") ~ "$30,000 to $39,999",
      variable %in% c("$40,000 to $44,999", "$45,000 to $49,999") ~ "$40,000 to $49,999",
      variable %in% c("$100,000 to $124,999", "$125,000 to $149,999", "$150,000 to $199,999", "$200,000 or more") ~ "$100,000 or more",
      TRUE ~ variable
    )
  ) %>%
  group_by(table_name, table_short_name, table_number, state, county, geography, GEOID, variable) %>%
  summarize(
    estimate = sum(estimate),
    moe = moe_sum(moe, estimate)
  ) %>%
  ungroup() %>%
  
  # convert to dot density
  tidycensus::as_dot_density(
    value = "estimate",
    values_per_dot = 20
  ) # convert MULTIPOLYGON to 20:1 POINT geometry

sf_household_income_rgv_pb_ep <- sf_acs_rgv_pb_ep %>%
  
  # get appropriate table rows
  filter(table_short_name == "Household income") %>% # pull specific ACS table
  filter(
    variable != "Total",
    !st_is_empty(.)
  ) %>% # filter out NA geometry to prep for dot density
  
  # consolidate variable groups
  mutate(
    variable = case_when(
      variable %in% c("Less than $10,000", "$10,000 to $14,999", "$15,000 to $19,999") ~ "Less than $20,000",
      variable %in% c("$20,000 to $24,999", "$25,000 to $29,999") ~ "$20,000 to $29,999",
      variable %in% c("$30,000 to $34,999", "$35,000 to $39,999") ~ "$30,000 to $39,999",
      variable %in% c("$40,000 to $44,999", "$45,000 to $49,999") ~ "$40,000 to $49,999",
      variable %in% c("$100,000 to $124,999", "$125,000 to $149,999", "$150,000 to $199,999", "$200,000 or more") ~ "$100,000 or more",
      TRUE ~ variable
    )
  ) %>%
  group_by(table_name, table_short_name, table_number, state, county, geography, GEOID, variable) %>%
  summarize(
    estimate = sum(estimate),
    moe = moe_sum(moe, estimate)
  ) %>%
  ungroup() %>%
  
  # convert to dot density
  tidycensus::as_dot_density(
    value = "estimate",
    values_per_dot = 20
  ) # convert MULTIPOLYGON to 20:1 POINT geometry

sf_household_income_ips <- sf_acs_ips %>%
  
  # get appropriate table rows
  filter(table_short_name == "Household income") %>% # pull specific ACS table
  filter(
    variable != "Total",
    !st_is_empty(.)
  ) %>% # filter out NA geometry to prep for dot density
  
  # consolidate variable groups
  mutate(
    variable = case_when(
      variable %in% c("Less than $10,000", "$10,000 to $14,999", "$15,000 to $19,999") ~ "Less than $20,000",
      variable %in% c("$20,000 to $24,999", "$25,000 to $29,999") ~ "$20,000 to $29,999",
      variable %in% c("$30,000 to $34,999", "$35,000 to $39,999") ~ "$30,000 to $39,999",
      variable %in% c("$40,000 to $44,999", "$45,000 to $49,999") ~ "$40,000 to $49,999",
      variable %in% c("$100,000 to $124,999", "$125,000 to $149,999", "$150,000 to $199,999", "$200,000 or more") ~ "$100,000 or more",
      TRUE ~ variable
    )
  ) %>%
  group_by(table_name, table_short_name, table_number, state, county, geography, GEOID, variable) %>%
  summarize(
    estimate = sum(estimate),
    moe = moe_sum(moe, estimate)
  ) %>%
  ungroup() %>%
  
  # convert to dot density
  tidycensus::as_dot_density(
    value = "estimate",
    values_per_dot = 20
  ) # convert MULTIPOLYGON to 20:1 POINT geometry

sf_household_income <- bind_rows(
  sf_household_income_houston,
  sf_household_income_tarrant,
  sf_household_income_ctx,
  sf_household_income_rgv_pb_ep,
  sf_household_income_ips
)


st_write_feather(sf_household_income, glue::glue("{working_dir}/sf_household_income_dd_2023.feather"))
```


```{r}
# students in poverty ----
# time 2410.67 mvp

sf_students_in_poverty_houston <- sf_acs_houston %>%
  
  # get appropriate table rows
  filter(table_short_name == "Students in poverty") %>% # pull specific ACS table
  filter(
    variable != "Total",
    !st_is_empty(.)
  ) %>% # filter out NA geometry to prep for dot density
  
  # convert to dot density
  tidycensus::as_dot_density(
    value = "estimate",
    values_per_dot = 20
  ) # convert MULTIPOLYGON to 20:1 POINT geometry

sf_students_in_poverty_tarrant <- sf_acs_tarrant %>%
  
  # get appropriate table rows
  filter(table_short_name == "Students in poverty") %>% # pull specific ACS table
  filter(
    variable != "Total",
    !st_is_empty(.)
  ) %>% # filter out NA geometry to prep for dot density
  
  # convert to dot density
  tidycensus::as_dot_density(
    value = "estimate",
    values_per_dot = 20
  ) # convert MULTIPOLYGON to 20:1 POINT geometry

sf_students_in_poverty_ctx <- sf_acs_ctx %>%
  
  # get appropriate table rows
  filter(table_short_name == "Students in poverty") %>% # pull specific ACS table
  filter(
    variable != "Total",
    !st_is_empty(.)
  ) %>% # filter out NA geometry to prep for dot density
  
  # convert to dot density
  tidycensus::as_dot_density(
    value = "estimate",
    values_per_dot = 20
  ) # convert MULTIPOLYGON to 20:1 POINT geometry

sf_students_in_poverty_rgv_pb_ep <- sf_acs_rgv_pb_ep %>%
  
  # get appropriate table rows
  filter(table_short_name == "Students in poverty") %>% # pull specific ACS table
  filter(
    variable != "Total",
    !st_is_empty(.)
  ) %>% # filter out NA geometry to prep for dot density
  
  # convert to dot density
  tidycensus::as_dot_density(
    value = "estimate",
    values_per_dot = 20
  ) # convert MULTIPOLYGON to 20:1 POINT geometry

sf_students_in_poverty_ips <- sf_acs_ips %>%
  
  # get appropriate table rows
  filter(table_short_name == "Students in poverty") %>% # pull specific ACS table
  filter(
    variable != "Total",
    !st_is_empty(.)
  ) %>% # filter out NA geometry to prep for dot density
  
  # convert to dot density
  tidycensus::as_dot_density(
    value = "estimate",
    values_per_dot = 20
  ) # convert MULTIPOLYGON to 20:1 POINT geometry

sf_students_in_poverty <- bind_rows(
  sf_students_in_poverty_houston,
  sf_students_in_poverty_tarrant,
  sf_students_in_poverty_ctx,
  sf_students_in_poverty_rgv_pb_ep,
  sf_students_in_poverty_ips
)


st_write_feather(sf_students_in_poverty, glue::glue("{working_dir}/sf_students_in_poverty_dd_2023.feather"))
```


```{r}
# households with children under 18 ----
# time: 2165.5 sec

sf_hh_children_under_18 <- sf_acs %>%
  
  # get appropriate table rows
  filter(table_short_name == "Households with children under 18") %>% # pull specific ACS table
  filter(!st_is_empty(.)) %>% # filter out NA geometry to prep for dot density
  
  # convert to dot density
  tidycensus::as_dot_density(
    value = "estimate",
    values_per_dot = 20
  ) # convert MULTIPOLYGON to 20:1 POINT geometry


# st_write_feather(sf_hh_children_under_18, glue::glue("{working_dir}/sf_hh_children_under_18_dd_2023.feather"))
```

```{r}
# mobility ----
# time: 412.84 sec

sf_mobility <- sf_acs %>%
  
  # get appropriate table rows
  filter(table_short_name == "Mobility") %>% # pull specific ACS table
  filter(!st_is_empty(.)) %>% # filter out NA geometry to prep for dot density
  
  # consolidate variable groups
  mutate(
    variable = case_when(
      variable %in% c("Moved within same county, 1 to 4 years", "Moved within same county, 5 to 17 years") ~ "Moved within same county",
      variable %in% c("Moved from different county within same state, 1 to 4 years", "Moved from different county within same state, 5 to 17 years") ~ "Moved from different county within same state",
      variable %in% c("Moved from different state, 1 to 4 years", "Moved from different state, 5 to 17 years") ~ "Moved from different state",
      variable %in% c("Moved from abroad, 1 to 4 years", "Moved from abroad, 5 to 17 years") ~ "Moved from abroad",
      TRUE ~ variable
    )
  ) %>%
  group_by(table_name, table_short_name, table_number, state, county, geography, GEOID, variable) %>%
  summarize(
    estimate = sum(estimate),
    moe = moe_sum(moe, estimate)
  ) %>%
  ungroup() %>%
  
  # convert to dot density
  tidycensus::as_dot_density(
    value = "estimate",
    values_per_dot = 20
  ) # convert MULTIPOLYGON to 20:1 POINT geometry


# st_write_feather(sf_mobility, glue::glue("{working_dir}/sf_mobility_dd_2023.feather"))
```


1. Filter out empty geographies
2. Then spatially join to isochrones

## Aggregate data at site + category level

1. Group by site location and ACS category
2. Summarize by sum() and moe_sum()

## Get dot densities

Now, we will bind all dot density data in one sf data frame.

```{r}
sf_dot_density_mvp <- bind_rows(
    sf_children_by_age_group_block_group,
    sf_children_by_age_group_tract,
    sf_children_in_poverty,
    sf_hh_children_under_18,
    sf_household_income,
    sf_mobility,
    sf_students_in_poverty
  ) %>%
  st_as_sf()
  
```


## Pin data

```{r pin_data}
# attach to pins board ----
board_posit <- board_connect()

board_posit %>%
  pin_write(
    sf_dot_density_mvp,
    name = "dashboard_dot_density_data",
    description = "Dot-density ACS data",
    type = "qs"
  )
```

